Compiling refinement types to dependent types

### 2020-07-20 tex/paper.tex
Build fix
Add \ctxok command to avoid code duplication

\newcommand{\unit}{u}
\newcommand{\Unit}{\textbf{Unit}}
\newcommand{\unitc}{\hat{u}}
\newcommand{\Unitc}{\widehat{\textbf{Unit}}}
\newcommand{\Int}{\textbf{Int}}

@@

The typing rules for the surface language are laid out in \figref{surface_typing}.
For the type well-formedness and term typing rules
we also assume each of them has a premise $\nonterm G \text{ ok}$
stating that the context is well-formed.

@@

\begin{figure}[ht]
  \footnotesize
  \drules[TCTX]{$\nonterm G \text{ ok}$}{context well-formedness}{Empty,Bind}
  \drules[TWF]{$\nonterm G \vdash \nonterm{ts}$}{type well-formedness}{TrueRef,Base,Conj,Arr,Rec}
  \drules[T]{$\nonterm G \vdash \nonterm{es} : \nonterm{ts}$}{term typing}{Unit,Var,Abs,App,Case,Con,Sub}
  \drules[ST]{$\nonterm G \vdash \nonterm{ts}_1 <: \nonterm{ts}_2$}{subtyping}{Base,Arr}
  \caption{Surface language typing}
  \label{fig:surface_typing}
\end{figure}

@@

\newcommand{\tranctx}{\mu_\Gamma}
\newcommand{\tranty}{\mu_\tau}
\newcommand{\tranterm}{\mu_\varepsilon}

@@

\paragraph{Contexts.}
We start with defining $\tranctx$,
which is the simplest translation of the three.
It merely maps the types of the bindings in a (well-formed) context
from the surface language into the core language.

\begin{equation}
\begin{aligned}
  & \tranctx(\emptyset)
      && \triangleq
    \emptyset                                               \\
  & \tranctx(\nonterm G, \gterm x : \nonterm{ts})
      && \triangleq
    \tranctx(\nonterm G), \gterm x : \tranty(\nonterm {G} \vdash \nonterm{ts})  \\
\end{aligned}
\end{equation}

@@

\begin{itemize}
  \item \ranchor{T-Unit}.
    \begin{align*}
      & \inferrule{\nonterm G \text{ ok}}{\nonterm G \vdash \unit : \{ \gterm v : \Unit | \top \}}
          \longmapsto
        \inferrule{ }{\nonterm{GC} \vdash \unitc : \Unitc}, \\
      \text{where}~&\nonterm{GC} = \tranctx(\nonterm G \text{ ok}).
    \end{align*}
    This rule has a single premise $\nonterm G \text{ ok}$ that the surface context is well-formed.
    We use $\Tranctx$ to translate the surface derivation of that premise into the core one,
    and extract the resulting context $\nonterm{GC}$.
    We then use that context for \ranchor{CT-UnitTerm}.
\end{itemize}

Having defined these functions, we formulate a simple but useful lemma:
\begin{lemma}[Translation totality]
  $\Tranctx$, $\Tranty$ and $\Tranterm$ are total.
\end{lemma}
\begin{proof}
  Firstly, each function is defined for all possible inputs.
  Then, each function recurses on structurally smaller inputs,
  and, since derivation trees are finite,
  this implies each function terminates for any input.
\end{proof}

>>>>

\newcommand{\unit}{u}
\newcommand{\Unit}{\textbf{Unit}}
\newcommand{\unitc}{\hat{u}}
\newcommand{\Unitc}{\widehat{\textbf{Unit}}}
\newcommand{\Int}{\textbf{Int}}
\newcommand{\ctxok}{\text{~ok}}

@@

The typing rules for the surface language are laid out in \figref{surface_typing}.
For the type well-formedness and term typing rules
we also assume each of them has a premise $\nonterm G \ctxok$
stating that the context is well-formed.

@@

\begin{figure}[ht]
  \footnotesize
  \drules[TCTX]{$\nonterm G \ctxok$}{context well-formedness}{Empty,Bind}
  \drules[TWF]{$\nonterm G \vdash \nonterm{ts}$}{type well-formedness}{TrueRef,Base,Conj,Arr,Rec}
  \drules[T]{$\nonterm G \vdash \nonterm{es} : \nonterm{ts}$}{term typing}{Unit,Var,Abs,App,Case,Con,Sub}
  \drules[ST]{$\nonterm G \vdash \nonterm{ts}_1 <: \nonterm{ts}_2$}{subtyping}{Base,Arr}
  \caption{Surface language typing}
  \label{fig:surface_typing}
\end{figure}

@@

\newcommand{\tranty}{\mu_\tau}
\newcommand{\tranterm}{\mu_\varepsilon}

@@

\paragraph{Contexts.}
As a warm-up,
we start with defining $\Tranctx$,
which is the simplest translation of the three.
It merely maps the types of the bindings in a (well-formed) context
from the surface language into the core language.

\begin{itemize}
  \item \ranchor{TCTX-Empty}.
    \begin{align*}
      & \inferrule{~}
                  {\emptyset \ctxok}
          \longmapsto
        \emptyset.
    \end{align*}
    The base case is trivial: we map the empty context to the empty context.
  \item \ranchor{TCTX-Bind}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\ \nonterm G \vdash \nonterm{ts}}
                  {\nonterm G , \gterm x : \nonterm{ts} \ctxok}
          \longmapsto
        \Tranctx(\nonterm G \ctxok), x : \tranty(\nonterm G \vdash \nonterm{ts}).
    \end{align*}

    We recurse on the prefix $\nonterm G$,
    which we are allowed to do due to the $\nonterm G \ctxok$ premise,
    and we use $\tranty$ to get the translated type of $\gterm x$,
    which we are also allowed to do due to the corresponding premise.
\end{itemize}

@@

\begin{itemize}
  \item \ranchor{T-Unit}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok}
                  {\nonterm G \vdash \unit : \{ \gterm v : \Unit | \top \}}
          \longmapsto
        \inferrule{~}
                  {\nonterm{GC} \vdash \unitc : \Unitc}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok).
    \end{align*}
    This rule has a single premise $\nonterm G \ctxok$ that the surface context is well-formed.
    We use $\Tranctx$ to translate the surface derivation of that premise into the core one,
    and extract the resulting context $\nonterm{GC}$.
    We then use that context for \ranchor{CT-UnitTerm}.
\end{itemize}

Having defined these functions, we formulate a simple but useful lemma:
\begin{lemma}[Translation totality]
  $\tranctx$, $\Tranty$ and $\Tranterm$ are total.
\end{lemma}
\begin{proof}
  Firstly, each function is defined for all possible inputs.
  Then, each function recurses on structurally smaller inputs,
  and, since derivation trees are finite,
  this implies each function terminates for any input.
\end{proof}
